### 2015-07-02 narrative/naming-service.tex
merged naming and discovery

\subsubsection{Naming Service}

A traditional POSIX naming service offers a hierarchical space consisting of
directories and files. This structure has led to scalability problems because
of the serialized access to a single source for creating and accessing files.
Several efforts~\cite{giga+,pvfs,others} have worked to reduce this contention
by doing things like reducing the serialized scope to a single directory or
subtree. While these approaches help, they do not address this key scalability
limitation.

Pure object stores, such as those popular in the big data
domain~\cite{memcached,others} avoid this bottleneck by strictly offering an
object ID with the application required to manage how this ID maps to something
of interest. This approach of removing the metadata service from the system
level completely can work well for scale out applications where data is created
or consumed by a single process at a time rather than potentially O(1 million)
processes all actively collectively for a single ``object''. To address this
case, having some system integrated metadata services to associate names with
these object is a preferable solution.

LWFS and Sirocco have take an approach similar to the pure object stores, but
with a focus on the HPC setting. They have abandonded a fully POSIX compliant
metadata service as the default model in favor of a
container/object/fork/addrress tuple for identifying data similar to those
used for pure object stores. By having a service that addresses the object
collections that comprise a single thing, such as a variable or timestep
output, just enough metadata is maintained to make the storage system usable
without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
namespace on the side kept in sync using a transaction process like
D2T~\cite{d2t} showing that this alternative approach can support traditional
POSIX API calls even though the underlying storage system uses a different
model.

The challenge this proposal brings is in the additional metadata required to
track all of the different pieces of a single variable across the storage
hierarchy both horizontally and vertically and keeping this metadata in sync
as the storage system (Sirocco) manages data according to user requests and
system load. Further, the profiled performance for annotating data according to
future read needs also must be tracked. Overall, since the end-to-end quality
of service requirements must drive all data placement and movement decisions,
the metadata and naming service must be rich enough, and low overhead enough,
to support these operations as well as maintaining data accessibility.

@@

@@

>>>>

\subsubsection{Naming and Discovery Service}

Given the well documented difficulties with POSIX-compliant naming and our
potentially heavily decomposed objects, we must take a new approach to naming.
Adding complexity is data migration based on system pressures. Our naming
service must support sufficient information to discover or track all variables
pieces across the entire storage hierarchy.

Hierarchical Storage Management (HSM) systems offer a strct caching approach
for managing different storage capacities trading off performance for capacity.
By maintaining a single namespace across all tiers, it is possible to list a
single directory view with files stored at different tiers. While this approach
to managing multi-tier storage works for whole files that fit on single logial
devices, it is far from ideal for scientific simulations.

Our goal with this proposal is to offer a similar capability, but use a finer
granularity and also manage discovering where the data may have migrated to
based on system pressures. Instead of a single file such as might be used to
store an entire timestep output for a simulation, we will demonstrate offering
the same capability at a subset of a single variable level. By shifting to a
finer-grained approach, we will enable more effective use of close/fast/small
storage tiers and manage the impact when there is insufficient space forcing
data to move. Traiditional HSM stores an entire file on a tier making room if
insufficient space is available. With this shift to a partial variable
granularity, a 1 PB output with 500 GB of ``high interest'' data can limit this
costly tier ussage to just 500 GB greatly enhancing usability. This reduces
pressure forcing data migration, but does not prevent it. We still need to be
able to find the data should it move based on system pressure.

The resulting naming and discovery service will attach the Sirocco
contaier/object/fork/address tuple with a special attribute indicating that
this record is of ``high interest'' strongly encouraging keeping the variable
in close/fast/small storage as long as possible. In addition to maintaining
this additional metadata, the application and/or middleware must provide this
information. The storage system simply needs to act appropriately on this data
based on the attribute.

The challenge for discovery is that potentially, data will migrate from where
it is initially stored to a new location within the storage system. Sirocco
offers an ability to search for data that has moved as well as forcing a
particular resilience-based replica be the ``authoritative'' version. We will
investigate if the current Sirocco functionality is capable of supporting the
new operating modes we wish to offer. Initial expectations suggest having
bounded time guarantees for finding data are critical for offering the quality
of service guarantees we wish to offer. Some potential approaches can build on
CRUSH~\cite{ceph} from Ceph to offer a map of where to search for data
sequentially. Given our multiple storage tiers, we would extend this using some
search limiting mechanism indicating to shift to the next tier to continue
searching because the requested data never made it to this location. This new
work will be an expansion of Sirocco's currently planned features.

The overriding theme for this proposal of providing a cooperative relationship
between the user and the storage system extends into this area as well. In
order for the user to negotiate with the storage system for data
quality/retrieval time tradeoff, additional data is required. Providing this
functionality requries the naming service be able to both return a list of data
versions, but also some indication of data quality and estimated retrieval time
for each. This estimate will be disucssed elsewhere in this proposal.

Structurally, the traditional POSIX naming service offeriing a hierarchical
space consisting of directories and files may be maintained for backwards
compatibility. However, this view will not offer the same strict semantics
POSIX requires. We must break these strict semantics to address scalability
problems because of the serialized access to a single source for creating and
accessing files.  Several efforts~\cite{giga+,pvfs,others} have worked to
reduce this contention by doing things like reducing the serialized scope to a
single directory or subtree. While these approaches help, they do not address
this key scalability limitation. Instead, we ar willing to offer a short
duration to eventually consistent POSIX view to offer the performance and
consistency requirements an exascale application demands.

While pure object stores, such as those popular in the big data
domain~\cite{memcached,others} avoid this bottleneck by strictly offering an
object ID with the application required to manage how this ID maps to something
of interest. This approach of removing the metadata service from the system
level completely can work well for scale out applications where data is created
or consumed by a single process at a time rather than potentially O(1 million)
processes all actively collectively for a single ``object''. To address this
case, having some system integrated metadata services to associate names with
these object is a preferable solution. While it will not have the same
synchronous consistency semantics as POSIX, we will offer something as close as
possible to address legacy application and command-line style data access and
migration to other storage types using these POSIX semantics.

We plan to address this scalability problem by continuing with the LWFS and
Sirocco model. LWFS and Sirocco have take an approach similar to the pure
object stores, but with a focus on the HPC setting. They have abandonded a
fully POSIX compliant metadata service as the default model in favor of a
container/object/fork/addrress tuple for identifying data similar to those used
for pure object stores. By having a service that addresses the object
collections that comprise a single thing, such as a variable or timestep
output, just enough metadata is maintained to make the storage system usable
without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
namespace on the side kept in sync using a transaction process like
D2T~\cite{d2t} showing that this alternative approach can support traditional
POSIX API calls even though the underlying storage system uses a different
model. Because we are not requiring the entire storage space be addressible
from a single tree root, we can offer multiple independent roots using the
Sirocco object storage. We will investigate how to make the naming and
discovery service scalable under the circumstances outlined above.

@@

While strongly encouraging data to stay in a particular type of storage tier,
ultimately it must be possible to flush this data to make room for other uses.
Sirocco offers a ``flush'' operation that forcibly migrates data to the
long-term resilience requirements freeing space in the scare resources. In
many, if not most cases, this will move data towards or actually on to a device
like tape intended for long-term storage.

@@

\begin{enumerate}

\item Milestone 1: Demonstrate a metadata service capable of serving both POSIX
clients and our clients, but without consideration for authentication, authorization, or data migration.

\item Milestone 2: Demonstrate a time bounded search approach for finding data
within the storage system to identify data the current location and retrieval
performance characteristics for data since migration. This will update the
cached information in the metadata service to short-circuit future requests for
the data presuming that it does not migrate again.

\item Milestone 3: Demonstrate incorporating A\&A and show scalability under
both application load as well as storage system pressures showing that we can
tolerate both loads and maintain quality of service.

\end{enumerate}
